---
title: "multiome_RNA_cell_populations"
author: "saraglbr"
output: html_document
---

Objective of this part of the code is to use the ATAC seq to identify populations of interest (EpiLCs and PGCLCs) based on regions of open chromatin 

#Libraries
```{r}
library(Seurat)
library(dplyr)
library(patchwork)
library(Signac)
library(SeuratDisk)
library(monocle3)
library(tidyr)
library(ggplot2)
library(org.Mm.eg.db)
library(clusterProfiler)
library(enrichplot)

```


#Load Seurat object after peak calling with MACS2 and normalization of both RNA and ATAC-seq 
```{r}
setwd("~/phd/scmultiome_analysis")
pbmc_filter <- LoadH5Seurat("multiome_SCT_PEAKS_only.h5seurat")
```



#Perform dimensionality reduction and genrate a UMAP for visualization
```{r fig.width=13, fig.height=6}
#Dimension reduction and SLI 
DefaultAssay(pbmc_filter) <- "peaks"
pbmc_filter <- FindTopFeatures(pbmc_filter, min.cutoff = 5)
pbmc_filter <- RunTFIDF(pbmc_filter)
pbmc_filter <- RunSVD(pbmc_filter)

#Correlation between dimensions and depth 
pdf("quality_ATAC.pdf")
DepthCor(pbmc_filter)
dev.off

#Dimensions 1 correlate with depth so it will be excluded from downstream analysis 
DefaultAssay(pbmc_filter) <- "peaks"
pbmc_filter <- FindNeighbors(object = pbmc_filter, reduction = 'lsi', dims = 2:15)
pbmc_filter <- FindClusters(object = pbmc_filter, verbose = FALSE, algorithm = 3, resolution= 0.35)
pbmc_filter <- RunUMAP(object = pbmc_filter, reduction = 'lsi', dims = 2:15)
pbmc_filter@meta.data$UMAPATAC <- pbmc_filter@reductions$lsi@cell.embeddings[, 1:2]

setwd("~/phd/thesis_figures/Single_Cell_Chapter/panels")
pdf("ATAC_UMAP.pdf")
DimPlot(object = pbmc_filter, label = TRUE,   cols = c('0'="#FDBF6F",'1'='#B3DE69', "2"="#33A02C", 
   '3'='#A4DFF2','4'='#4B4BF7', "5" ="#6A3D9A", '6'= "#FB9A99", 
     "8"= '#AC8F14', "7" = "#E31A1C", "9"="#FF7F00"
   ) ) 
dev.off
```

```{r fig.width=13, fig.height=6}
setwd("~/phd/thesis_figures/Single_Cell_Chapter/panels")
pdf("ATAC_UMAP_after_filtering.pdf")
FeaturePlot(pbmc_filter, features = c("nCount_ATAC", 'TSS.enrichment', 'blacklist_ratio', 'nucleosome_signal'), ncol = 2, cols = c("green", "red"))
dev.off
```



#All - Comapirson between ATAC-seq of all EpiLCs vs all day 2 PGCLC aggregates
```{r}
pbmc_filter
DimPlot(pbmc_filter, reduction="umap")
DefaultAssay(pbmc_filter) <- "peaks"

pbmc_filter <- FindClusters(pbmc_filter, verbose=FALSE, algorithm = 3, resolution=0.05)
DimPlot(pbmc_filter, reduction="umap")

ATAC_epi <- FindMarkers(pbmc_filter,
            ident.1 = "0")

ATAC_epi <- ATAC_epi[ATAC_epi$p_val<0.1&ATAC_epi$avg_log2FC>0.1, ]
open_epi_genes <- ClosestFeature(pbmc_filter, regions = rownames(ATAC_epi))


ATAC_d2 <- FindMarkers(pbmc_filter, 
                       ident.1="1")

ATAC_d2 <- ATAC_d2[ATAC_d2$p_val<0.05&ATAC_d2$avg_log2FC>0.25, ]
open_D2_genes <- ClosestFeature(pbmc_filter, regions=rownames(ATAC_d2))

#GO enrichemnt 
#Epi
FeaturePlot(pbmc_filter, features = rownames(ATAC_epi[1, ]),
            pt.size=0.1,
            cols= c("burlywood1", "darkblue"))

entrez_epi <- bitr(open_epi_genes[ ,"gene_name"], fromType="SYMBOL", 
                   toType = "ENTREZID", OrgDb = org.Mm.eg.db)
epi_enrichment <- enrichGO(entrez_epi$ENTREZID,
                           OrgDb = org.Mm.eg.db,readable = T,
                           ont = "BP", pvalueCutoff = 0.05, qvalueCutoff = 0.10)
epi_enrichment_plot <- barplot(epi_enrichment, showCategory=5)
epi_enrichment_plot
epi_enrichment <- epi_enrichment@result

#D2
FeaturePlot(pbmc_filter, features = rownames(ATAC_d2[1, ]),
            pt.size=0.1,
            cols= c("burlywood1", "darkblue"))

entrez_d2 <- bitr(open_D2_genes[ ,"gene_name"], fromType="SYMBOL", 
                   toType = "ENTREZID", OrgDb = org.Mm.eg.db)
d2_enrichment <- enrichGO(entrez_d2$ENTREZID,
                           OrgDb = org.Mm.eg.db,readable = T,
                           ont = "BP", pvalueCutoff = 0.05, qvalueCutoff = 0.10)
d2_enrichment_plot <- barplot(d2_enrichment, showCategory=5)
d2_enrichment_plot
d2_enrichment <- d2_enrichment@result



epi_enrichment <- merge(epi_enrichment, open_epi_genes )
setwd("~/phd/thesis_figures/Single_Cell_Chapter/multiome/between_samples")
write.csv(epi_enrichment, "Epi_enrichment.csv")
write.csv(ATAC_epi, "ATAC_epi_open.csv")

write.csv(d2_enrichment, "D2_enrichment.csv")
write.csv(ATAC_d2, "ATAC_d2_open.csv")



setwd("~/phd/thesis_figures/Single_Cell_Chapter/multiome/between_samples")
pdf("enrichment_plot_open_epi.pdf", height=4, width=5)
epi_enrichment_plot
dev.off
d2_enrichment_plot
pdf("essrb_atac.pdf")
plot1 | plot2
dev.off

```




#Annotate clusters by assessing the differentially accesible region and their function

```{r}
#Create a function to find differentially accesible regions for each cluster, save markers into a CSV and return a list of the top 20 genes  
analyze_cluster <- function(cluster_ident) {
  # Find markers
  cluster_markers <- FindMarkers(pbmc_filter, ident.1 = cluster_ident, min.pct = 0.25)

  # Save cluster markers to CSV
  cluster_markers_df <- as.data.frame(cluster_markers)
  write.csv(cluster_markers_df, file = paste0("ATAC_cluster", cluster_ident, "_markers.csv"), row.names = TRUE)
  
  # Return the top 20 genes
  return(cluster_markers[cluster_markers$avg_log2FC>0.025, ])
}

```

```{r}
#Run the fuction for all 9 clusters
cluster0 <- analyze_cluster(0)
cluster1 <- analyze_cluster(1)
cluster2 <- analyze_cluster(2)
cluster3 <- analyze_cluster(3)
cluster4 <- analyze_cluster(4)
cluster5 <- analyze_cluster(5)
cluster6 <- analyze_cluster(6)
cluster7 <- analyze_cluster(7)
cluster8 <- analyze_cluster(8)
cluster10 <- analyze_cluster(9)
```



#Focus on EpiLC clusters 
```{r}
#Load differentially accesbile regions (DAR)
setwd("~/phd/thesis_figures/Single_Cell_Chapter/multiome/ATAC_clusters_markers")
cluster0 <- read.csv("ATAC_cluster0_markers.csv")
cluster2 <- read.csv("ATAC_cluster2_markers.csv")
cluster3 <- read.csv("ATAC_cluster3_markers.csv")
cluster5 <- read.csv("ATAC_cluster5_markers.csv")

#Find the closest genes to each DAR
cluster0_Genes <- ClosestFeature(pbmc_filter, regions = rownames(cluster0))
cluster2_Genes <- ClosestFeature(pbmc_filter, regions = rownames(cluster2))
cluster3_Genes <- ClosestFeature(pbmc_filter, regions = rownames(cluster3))
cluster5_Genes <- ClosestFeature(pbmc_filter, regions = rownames(cluster5))


#Attempt to do GO enrichment with differentially open genes -  no result
genes_open <- cluster5_Genes[ ,"gene_name"] 
open_entrez = bitr(genes_open, fromType = "SYMBOL", toType = c("ENTREZID"), OrgDb =
                     org.Mm.eg.db)  
open_enrichment = enrichGO(gene = open_entrez$ENTREZID, OrgDb = org.Mm.eg.db,readable = T,
                           ont = "BP", pvalueCutoff = 0.05, qvalueCutoff = 0.10)
open_enrichment_plot = barplot(open_enrichment, showCategory=5, xlab="number of genes invovled")
open_enrichment_plot
open_go <- open_enrichment@result

#Visualize top regions for each cluster
pdf("freature_maps_ATACEPI_markers.pdf", height=6.5, width=8)
FeaturePlot(
  object = pbmc_filter,
  features =   c("chr5-112031473-112032606", #cluster0
                 "chr1-167497153-167498839", #cluster2
                 "chr15-75085448-75087030", #cluster3
                 "chr2-144247429-144248184"), #cluster5
  cols = c("burlywood1", "darkblue"))

#Visualize regions close to relevant genes 
FeaturePlot(
  object = pbmc_filter,
  features =   c("chr2-153688135-153689726", #dnmt3b
                  "chr5-19565430-19566988", # Magi2 - ovaries 
                "chr5-98283593-98284684", #Fgf5
                "chr11-35294716-35295587" #slit3
                ), 
  cols = c("burlywood1", "darkblue"))
dev.off

```




#Select only EpiLCs to increase resoltuion 
```{r}
pbmc_epi <- subset(pbmc_filter, subset = sample == "EpiLC")

ranges <- pbmc_epi@assays[["peaks"]]@ranges
setwd("~/phd/thesis_figures/Single_Cell_Chapter/multiome/EpiLCs")
write.csv(ranges, "all_peaks_epi.csv")
```

```{r}
#------------------------------------UMAP epi ATAC only---------------------------------
DefaultAssay(pbmc_epi) <- "peaks" #Dims 1-5 correlate with depth
quality <- DepthCor(pbmc_epi) 
pbmc_epi <- FindNeighbors(object = pbmc_epi, reduction = 'lsi', dims = 3:14)
pbmc_epi <- FindClusters(object = pbmc_epi, verbose = FALSE, algorithm = 3, resolution= 0.2)
pbmc_epi <- RunUMAP(object = pbmc_epi, reduction = 'lsi', dims = 3:14)


ATAC_epi <- DimPlot(object = pbmc_epi, reduction = "umap", 
                  cols = c('0'="#FDBF6F",'1'="#33A02C", 
   '2'='#A4DFF2', "3" ="#6A3D9A", '6'= "#FB9A99", 
     "8"= '#AC8F14', "7" = "#E31A1C", "9"="#FF7F00"
   
   ) ) 

setwd("~/phd/thesis_figures/Single_Cell_Chapter/panels/EpiLCs")
pdf("UMAP_EpiLC_ATAC.pdf", height=6, width=8)
quality
ATAC_epi
dev.off
```

```{r}

#--Find differentially open regions between Clusters 4 & 5(most different)------------
da_peaks <- FindMarkers(
  pbmc_epi,
  ident.1 = "1")

#Plot difference of top differentially open region 
plot1 <- VlnPlot(
  object = pbmc_epi,
  features = rownames(da_peaks)[1],
  pt.size = 0.05,
    cols = c('0'="#FDBF6F",'1'="#33A02C", 
   '2'='#A4DFF2', "3" ="#6A3D9A", '6'= "#FB9A99", 
     "8"= '#AC8F14', "7" = "#E31A1C", "9"="#FF7F00"
   
   ) )

plot2 <- FeaturePlot(
  object = pbmc_epi,
  features = rownames(da_peaks)[1],
  pt.size = 0.2,
  cols = c("burlywood1", "darkblue")
)

setwd("~/phd/thesis_figures/Single_Cell_Chapter/panels/EpiLCs")
pdf("EpiLC_top_DAR.pdf", height=6, width=8)
plot1 | plot2
dev.off

```

```{r}

#Sort into more open/close lists
open <- rownames(da_peaks[da_peaks$avg_log2FC > 0.25, ])
close<- rownames(da_peaks[da_peaks$avg_log2FC < -0.2, ])

#Find closest gene
da_peaks_Genes <- ClosestFeature(pbmc_epi, regions = open)
da_peaks_Genes_closed <- ClosestFeature(pbmc_epi, regions = close)


#Attempt to do GO enrichment with differentially open genes
genes_open <- da_peaks_Genes[ ,"gene_name"] # open in PGCLCs
open_entrez = bitr(genes_open, fromType = "SYMBOL", toType = c("ENTREZID"), OrgDb =
                     org.Mm.eg.db)  
open_enrichment = enrichGO(gene = open_entrez$ENTREZID, OrgDb = org.Mm.eg.db,readable = T,
                           ont = "BP", pvalueCutoff = 0.05, qvalueCutoff = 0.10)
open_enrichment_plot = barplot(open_enrichment, showCategory=5, xlab="number of genes invovled")
open_enrichment_plot
open_go <- open_enrichment@result



#Save tables as csv
setwd("~/phd/thesis_figures/Single_Cell_Chapter/panels/EpiLCs")
write.csv(da_peaks, "DARs_cluster3.csv")
write.csv(da_peaks_Genes, "Open_3_genes.csv")
write.csv(da_peaks_Genes_closed, "Closed_4vs3_genes.csv")

```



```{r}
#------------Find differentially open regions between Clusters 4 and everything else------------
DefaultAssay(pbmc_epi) <- 'peaks'
da_peaks <- FindMarkers(
  object = pbmc_epi,
  ident.1 = "1",
  ident.2 = "2",
  test.use = 'LR',
  latent.vars = 'nCount_peaks'
)

head(da_peaks)

fc <- FoldChange(pbmc_epi, ident.1 = "1", ident.2 = "2")


#Plot difference of top differentially open region 
plot1 <- VlnPlot(
  object = pbmc_epi,
  features = "chr7-126491872-126493023",
  pt.size = 0.1
)
plot2 <- FeaturePlot(
  object = pbmc_epi,
  features =  "chr7-126491872-126493023",
  pt.size = 0.1,
  cols = c("goldenrod1", "royalblue4")
)
plot1 | plot2


setwd("~/phd/thesis_figures/Single_Cell_Chapter/panels/EpiLCs")
pdf("EpiLC_top_DAR_C4.pdf", height=6, width=8)
plot1 | plot2
dev.off


#Sort into more open/close lists
open <- rownames(da_peaks[da_peaks$avg_log2FC > 0.2, ])

#Find closest gene
da_peaks_Genes <- ClosestFeature(pbmc_epi, regions = open)


setwd("~/phd/thesis_figures/Single_Cell_Chapter/panels/EpiLCs")
write.csv(da_peaks, "REGIONS_cluster4.csv")

```


```{r fig.height=2.5, fig.width=9}
#---------------------------Divide cells into Otx2 pos and neg-----------------------------
DefaultAssay(pbmc_epi) <- "SCT"

setwd("~/phd/thesis_figures/Single_Cell_Chapter/panels/EpiLCs")
pdf("EpiLC_marker_expression_ATAC_umap.pdf", height=2.5, width=13)
FeaturePlot(pbmc_epi, features = c("Otx2", "Fgf5", "Dnmt3b", "Nanog"),
            cols = c("cadetblue2", "red"), reduction = "umap", ncol=4)
dev.off


poscells <- WhichCells(pbmc_epi, expression = Otx2 > 0)
negcells <- WhichCells(pbmc_epi, expression = Otx2 == 0)
pbmc_epi$Otx2_expression<- ifelse(colnames(pbmc_epi) %in% poscells, "Pos", "Neg")
DimPlot(object = pbmc_epi,  reduction = "umap", group.by = "Otx2_expression") 

#Find differentially open regions between Otx2 positive and Otx2 negative------------
pbmc_epi <- RegroupIdents(pbmc_epi, metadata = "Otx2_expression")
Idents(object = pbmc_epi) <- pbmc_epi@meta.data$Otx2_expression
DefaultAssay(pbmc_epi) <- "peaks" 
da_peaks_otx2 <- FindMarkers(
  pbmc_epi,
  ident.1 = "Pos",
  ident.2 = "Neg") 

#Nothing found, try again with student t-test
da_peaks_otx2 <- FindMarkers(
  pbmc_epi,
  ident.1 = "Pos",
  ident.2 = "Neg",
  test.use = "roc") 

#Warning: No features pass logfc.threshold threshold; returning empty data.frame

```


#Focus on ATAC-seq clusters in day 2 PGCLC aggregates
```{r width=10}
setwd("~/phd/thesis_figures/Single_Cell_Chapter/multiome/ATAC_clusters_markers")
cluster1 <- read.csv("ATAC_cluster1_markers.csv")
cluster4 <- read.csv("ATAC_cluster4_markers.csv")
cluster6 <- read.csv("ATAC_cluster6_markers.csv")
cluster7 <- read.csv("ATAC_cluster7_markers.csv")
cluster8 <- read.csv("ATAC_cluster8_markers.csv")
cluster9 <- read.csv("ATAC_cluster9_markers.csv")


cluster1_Genes <- ClosestFeature(pbmc_filter, regions = cluster1$X)
cluster4_Genes <- ClosestFeature(pbmc_filter, regions = cluster4$X)
cluster6_Genes <- ClosestFeature(pbmc_filter, regions = cluster6$X)
cluster7_Genes <- ClosestFeature(pbmc_filter, regions = cluster7$X)
cluster8_Genes <- ClosestFeature(pbmc_filter, regions = cluster8$X)
cluster9_Genes <- ClosestFeature(pbmc_filter, regions = cluster9$X)


```


```{r height=10, width=15, }
#Visualize top differentially accesible regions - related to relevant genes 
setwd("~/phd/thesis_figures/Single_Cell_Chapter/multiome/ATAC_clusters_markers")
pdf("feature_DAY2_markers.pdf", height=6, width=10)
FeaturePlot(
  object = pbmc_filter,
  features =   c("chr7-79789160-79789795", #cluster1 Mesp1, mesoderm
                 "chr16-22799816-22801082", #Crygs
                 "chr4-141292846-141293433", #Epha2 related to primitive streak "EphA2 expression is first observed along the primitive streak in gastrulation stage embryos (6.5–7.5 days post coitum (dpc))." cluster6 (pink)
                 "chr12-86501635-86502146", #essrb (PGCLCs)
                 "chr16-45124454-45125157", #Ccdc80
                 "chr5-128541394-128542551" #5930412G12Rik
                 ),
   pt.size = 0.0001, cols = c("burlywood1", "darkblue"), ncol=3)
dev.off

```































```{r}

#Sort into more open/close lists
open <- rownames(da_peaks[da_peaks$avg_log2FC > 0.25, ])
close<- rownames(da_peaks[da_peaks$avg_log2FC < -0.2, ])

#Find closest gene
da_peaks_Genes <- ClosestFeature(pbmc_epi, regions = open)
da_peaks_Genes_closed <- ClosestFeature(pbmc_epi, regions = close)


#Attempt to do GO enrichment with differentially open genes
genes_open <- da_peaks_Genes[ ,"gene_name"] # open in PGCLCs
open_entrez = bitr(genes_open, fromType = "SYMBOL", toType = c("ENTREZID"), OrgDb =
                     org.Mm.eg.db)  
open_enrichment = enrichGO(gene = open_entrez$ENTREZID, OrgDb = org.Mm.eg.db,readable = T,
                           ont = "BP", pvalueCutoff = 0.05, qvalueCutoff = 0.10)
open_enrichment_plot = barplot(open_enrichment, showCategory=5, xlab="number of genes invovled")
open_enrichment_plot
open_go <- open_enrichment@result



#Save tables as csv
setwd("~/phd/thesis_figures/Single_Cell_Chapter/panels/EpiLCs")
write.csv(da_peaks, "DARs_cluster3.csv")
write.csv(da_peaks_Genes, "Open_3_genes.csv")
write.csv(da_peaks_Genes_closed, "Closed_4vs3_genes.csv")

```


```{r fig.height=2.5, fig.width=9}
#---------------------------Divide cells into Otx2 pos and neg-----------------------------
DefaultAssay(pbmc_epi) <- "SCT"

setwd("~/phd/thesis_figures/Single_Cell_Chapter/panels/EpiLCs")
pdf("EpiLC_marker_expression_ATAC_umap.pdf", height=2.5, width=13)
FeaturePlot(pbmc_epi, features = c("Otx2", "Fgf5", "Dnmt3b", "Nanog"),
            cols = c("cadetblue2", "red"), reduction = "umap", ncol=4)
dev.off


poscells <- WhichCells(pbmc_epi, expression = Otx2 > 0)
negcells <- WhichCells(pbmc_epi, expression = Otx2 == 0)
pbmc_epi$Otx2_expression<- ifelse(colnames(pbmc_epi) %in% poscells, "Pos", "Neg")
DimPlot(object = pbmc_epi,  reduction = "umap", group.by = "Otx2_expression") 

#Find differentially open regions between Otx2 positive and Otx2 negative------------
pbmc_epi <- RegroupIdents(pbmc_epi, metadata = "Otx2_expression")
Idents(object = pbmc_epi) <- pbmc_epi@meta.data$Otx2_expression
DefaultAssay(pbmc_epi) <- "peaks" 
da_peaks_otx2 <- FindMarkers(
  pbmc_epi,
  ident.1 = "Pos",
  ident.2 = "Neg") 

#Nothing found, try again with student t-test
da_peaks_otx2 <- FindMarkers(
  pbmc_epi,
  ident.1 = "Pos",
  ident.2 = "Neg",
  test.use = "roc") 

#Warning: No features pass logfc.threshold threshold; returning empty data.frame

```

```{r fig.height=2.5, fig.width=9}
#UMAP - 10dim 0.5res
DefaultAssay(pbmc_epi) <- "SCT"
pbmc_epi <- FindNeighbors(pbmc_epi, dims = 1:10)
pbmc_epi <- FindClusters(pbmc_epi, resolution = 0.2)
pbmc_epi <- RunUMAP(pbmc_epi, dims = 1:10)

#UMAP coloured by clusters
RNA_epi <- DimPlot(pbmc_epi, reduction = "umap")

#Color UMAP based on sample
setwd("~/phd/thesis_figures/Single_Cell_Chapter/panels/EpiLCs")
pdf("Epi_UMAPs.pdf", height=6, width=5.5)
RNA_epi
ATAC_epi
dev.off


setwd("~/phd/thesis_figures/Single_Cell_Chapter/panels/EpiLCs")
pdf("EpiLC_marker_expression_RNA_umap.pdf", height=2.5, width=13)
FeaturePlot(pbmc_epi, features = c("Mecom", "Fgf5", "Dnmt3b", "Nanog"),
            cols = c("cadetblue2", "red"), reduction = "umap", ncol=4)
dev.off
```
```{r}
setwd("~/phd/thesis_figures/Single_Cell_Chapter/panels")
SaveH5Seurat(pbmc_epi, filename ="multiome_Epi", overwrite = TRUE)
```
```{r}
setwd("~/phd/thesis_figures/Single_Cell_Chapter/panels")
pbmc_epi <- LoadH5Seurat("multiome_Epi.h5seurat")

```


```{r}
setwd("~/phd/thesis_figures/Single_Cell_Chapter/panels/EpiLCs")
analyze_cluster <- function(cluster_ident) {
  # Find markers
  cluster_markers <- FindMarkers(pbmc_epi, ident.1 = cluster_ident, min.pct = 0.25)
  
  # Save cluster markers to CSV
  cluster_markers_df <- as.data.frame(cluster_markers)
  write.csv(cluster_markers_df, file = paste0("EPI_cluster", cluster_ident, "_markers.csv"), row.names = TRUE)
  
  # Return the top 20 genes
  return(rownames(head(cluster_markers, n = 20)))
}

#Run the fuction for all 10 clusters
cluster0 <- analyze_cluster(0)
cluster1 <- analyze_cluster(1)
cluster2 <- analyze_cluster(2)
cluster3 <- analyze_cluster(3)
cluster4 <- analyze_cluster(4)


cluster0
cluster1 
cluster2 
cluster3 
cluster4 


```
```{r fig.height=2.5, fig.width=9}
setwd("~/phd/thesis_figures/Single_Cell_Chapter/panels/EpiLCs")
pdf("EpiLC_marker_expression_RNA_umap.pdf", height=2.5, width=10.5)
FeaturePlot(pbmc_epi, features = c(
  "Abcc4","Arap2","Flna", "mt-Atp8"
),
            cols = c("cadetblue2", "red"), reduction = "umap", ncol=4)
dev.off

FeaturePlot(pbmc_epi, features = c(
  "Abcc4", "Peg3", 
  "Ano3", "Arap2",
  "Mecom", "Pleg",
  "Anks1b", "Rhox5",
  "Flna"
),
            cols = c("cadetblue2", "red"), reduction = "umap", ncol=4)


```
#Select PGCLCs
```{r}
pbmc_filter$sample[pbmc_filter$sample=="NA"] <- "D2 aggregate"
table(pbmc_filter@meta.data$sample)

pbmc_pg <- subset(pbmc_filter, subset = sample ==  "D2 aggregate")

```

```{r}
VlnPlot(
  object = pbmc_filter,
  features = c("nCount_ATAC", 'TSS.enrichment', 'blacklist_ratio', 'nucleosome_signal'),
  pt.size = 0,
  ncol = 4,
  group.by="sample"
  )
```
```{r}
DefaultAssay(pbmc_pg) <- "SCT"

#PCA
pbmc_pg <- RunPCA(pbmc_pg)
ElbowPlot(pbmc_pg)

DimHeatmap((pbmc_pg), dims = 5:11, cells = 500, balanced = TRUE)

pbmc_pg <- FindNeighbors(pbmc_pg, dims = 2:10)
pbmc_pg <- FindClusters(pbmc_pg, resolution = 0.25)
pbmc_pg <- RunUMAP(pbmc_pg, dims = 2:10)
RNA_PGCLC_UMAP <- DimPlot(object = pbmc_pg, reduction = "umap", label = FALSE, cols =  c("#D39200", "#93AA00" , "brown1", "#00B9E3", "#DB72FB", "#FF61C3")) 



setwd("~/phd/thesis_figures/Single_Cell_Chapter/multiome/PGCLC_on_its_own")
pdf("UMAP_PGCLC.pdf", height=6, width=7)
RNA_PGCLC_UMAP
dev.off

FeaturePlot(pbmc_pg, features = c("Esrrb","Tfap2c", 
                                      "T", "Mixl1", "Krt19",
                                      "Otx2"),
            cols = c("cadetblue2", "red"), reduction = "umap", ncol=3)


FeaturePlot(pbmc_pg, features = c("Lef1","Tcf7l1", 
                                      "Zeb2", "Grhl2"),
            cols = c("cadetblue2", "red"), reduction = "umap", ncol=2)


```

```{r}
#------------------------------------UMAP epi ATAC only---------------------------------
DefaultAssay(pbmc_pg) <- "peaks" #Dims 1-5 correlate with depth
DepthCor(pbmc_pg) 
pbmc_pg <- FindNeighbors(object = pbmc_pg, reduction = 'lsi', dims = 2:18)
pbmc_pg <- FindClusters(object = pbmc_pg, verbose = FALSE, algorithm = 3, resolution= 0.25)
pbmc_pg <- RunUMAP(object = pbmc_pg, reduction = 'lsi', dims = 2:18)




FeaturePlot(pbmc_pg, features = c("Lef1","Tcf7l1", 
                                      "Zeb2", "Grhl2"),
            cols = c("cadetblue2", "red"), reduction = "umap", ncol=2)

UMAP_ATAC <- DimPlot(object = pbmc_pg, reduction = "umap",
       cols = c('0'='#B3DE69', "1"="#4B4BF7", 
   '2'='#FB9A99','3'='#E31A1C', "4" ="#AC8F14", 
   "5" ="#FF7F00"),
   label=FALSE,  pt.size=0.9) 


FeaturePlot(pbmc_pg, features = c("Esrrb","Tfap2c", 
                                      "T", "Mixl1", "Fgf8",
                                      "Zeb2"),
            cols = c("cadetblue2", "red"), reduction = "umap", ncol=3)



FeaturePlot(pbmc_pg, features = c("nCount_ATAC", 'TSS.enrichment', 'blacklist_ratio', 'nucleosome_signal'), ncol = 2, cols = c("green", "red"))

new.cluster.ids <- c("Mesoderm", "Intermedaite_1", 
                     "Intermediate_2", "PGCLCs", "Epithelia-like",
                      "Vascular-like")
names(new.cluster.ids) <- levels(pbmc_pg)
pbmc_pg <- RenameIdents(pbmc_pg, new.cluster.ids)
DimPlot(pbmc_pg, reduction = "umap", label = TRUE, label.box=TRUE) + NoLegend()





RNA_clusters_UMAP_ATAC <- DimPlot(pbmc_pg, reduction = "umap", label = TRUE, label.box=TRUE) + NoLegend()


setwd("~/phd/thesis_figures/Single_Cell_Chapter/multiome/PGCLC_on_its_own")
pdf("UMAP_PGCLC_ATAC.pdf", height=6, width=6)
UMAP_ATAC
dev.off



```
```{r}

#Create a function to find markers for each cluster, and save markers into a CSV, returns a list of the top 20 genes  
analyze_cluster <- function(cluster_ident) {
  # Find markers
  cluster_markers <- FindMarkers(pbmc_pg, ident.1 = cluster_ident, min.pct = 0.25)
  
  # Save cluster markers to CSV
  cluster_markers_df <- as.data.frame(cluster_markers)
  write.csv(cluster_markers_df, file = paste0("cluster", cluster_ident, "_markers.csv"), row.names = TRUE)
  
  # Return the top 20 genes
  return(cluster0 <- cluster_markers[cluster_markers$p_val<0.05
                         &
                           cluster_markers$avg_log2FC> 0.25 ,  ])
}

```

```{r}
#Run the function for all day 2 PGCLC aggregte clusters 
setwd("~/phd/thesis_figures/Single_Cell_Chapter/multiome/ATAC_clusters_markers/pgclc_only")
cluster0 <- analyze_cluster(0)
cluster1 <- analyze_cluster(1)
cluster2 <- analyze_cluster(2)
cluster3 <- analyze_cluster(3)
cluster4 <- analyze_cluster(4)
cluster5 <- analyze_cluster(5)

#Find closest feature 
cluster0_genes<- ClosestFeature(pbmc_pg, regions = rownames(cluster0))
cluster1_genes<- ClosestFeature(pbmc_pg, regions = rownames(cluster1))
cluster2_genes<- ClosestFeature(pbmc_pg, regions = rownames(cluster2))
cluster3_genes<- ClosestFeature(pbmc_pg, regions = rownames(cluster3))
cluster4_genes<- ClosestFeature(pbmc_pg, regions = rownames(cluster4))
cluster5_genes<- ClosestFeature(pbmc_pg, regions = rownames(cluster5))
```

```{r}
#annotation of peaks - annotation done with HOMER here just check the frequency of region type
d2_annotation <- read.csv("~/phd/thesis_figures/Single_Cell_Chapter/multiome/between_samples/D2_peak_annotation.csv")
epi_annotation <- read.csv("~/phd/thesis_figures/Single_Cell_Chapter/multiome/between_samples/epi_peak_annotation.csv")
d2_annotation$Annotation <- gsub("\\s*\\(.*?\\)\\s*", "", d2_annotation$Annotation)
epi_annotation$Annotation <- gsub("\\s*\\(.*?\\)\\s*", "", epi_annotation$Annotation)
d2_annotation_count <- as.data.frame(table(d2_annotation$Annotation))
epi_annotation_count <- as.data.frame(table(epi_annotation$Annotation))
ggplot(d2_annotation_count, aes(x="", y=Freq, fill=Var1)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0)
ggplot(epi_annotation_count, aes(x="", y=Freq, fill=Var1)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0)
```

```{r}
#GO enrichment of all clusters
go_save <- function(genes){
  genes_open <- genes[ ,"gene_name"] # open in PGCLCs
open_entrez = bitr(genes_open, fromType = "SYMBOL", toType = c("ENTREZID"), OrgDb =
                     org.Mm.eg.db)  
open_enrichment = enrichGO(gene = open_entrez$ENTREZID, OrgDb = org.Mm.eg.db,readable = T,
                           ont = "BP", pvalueCutoff = 0.05, qvalueCutoff = 0.10)
open_enrichment_plot = barplot(open_enrichment, showCategory=5, xlab="number of genes invovled")
open_go <- open_enrichment@result
name <- paste("~/phd/thesis_figures/Single_Cell_Chapter/multiome/ATAC_clusters_markers/", deparse(substitute(genes)), "GO.csv", sep="")
write.csv(open_go, name)
return(open_enrichment_plot)
}

go_save(cluster1_Genes)
go_save(cluster4_Genes)
go_save(cluster6_Genes)
go_save(cluster7_Genes)
go_save(cluster8_Genes)
go_save(cluster9_Genes)

```
