---
title: "extra QC and normalization"
output: html_notebook
---

```{r}
library(Seurat)
library(dplyr)
library(ggplot2)
library(Signac)
```

#After regressing cell cycle and mitochondrial reads cluster analysis was done again but one of the EpiLC clusters didn't match previous analysis, looked into it and it was mostly made up by low-quality cells, so did some QC re-filtering to get rid of them. 

then normalize & re-run UMAPS

```{r}
#Visualize UMAP
multiome <- FindNeighbors(multiome, dims = 1:17)
multiome <- FindClusters(multiome, resolution = 0.37)
multiome <- RunUMAP(multiome, dims = 1:17, reduction = "pca", seed.use=30)

DimPlot(multiome, reduction = "umap", label =TRUE)
DimPlot(multiome, reduction = "umap", group.by="sample")
DimPlot(multiome, reduction = "umap", group.by="Phase")

```
```{r}
#One of the markers doesn't match previous analysis, check markers
cluster_markers <- FindMarkers(multiome, ident.1 = "6", min.pct = 0.25)

FeaturePlot(multiome, reduction ="umap", features= c("Malat1", "Prim2", "Plekgh1"))

#Malat1 (top marker) can be associated with dying cells, check quality metric per cluster check if they are affecting UMAP

FeaturePlot(multiome, reduction ="umap", features= c("nCount_RNA", "nFeature_RNA"))
VlnPlot(multiome, features="nCount_RNA") #Clusters 0 and 6 significantly lower than the rest
VlnPlot(multiome, features="nFeature_RNA") # same issue same clusters
VlnPlot(multiome, features="percent.mt") #no difference per cluster
VlnPlot(multiome, features="percent.mt", y.max=1) #no difference per cluster

```

##Re-filter cells 
```{r}
#Visualize different cut offs on the plot 

#ncount
VlnPlot(multiome, features="nCount_RNA") + geom_hline(yintercept = 5100, color = "black", size=1)
VlnPlot(multiome, features= c("nCount_RNA"), group.by="sample") + geom_hline(yintercept = 5100, color = "black", size=1)

#nfeature
VlnPlot(multiome, features="nFeature_RNA") + geom_hline(yintercept = 2300, color = "black", size=1)
VlnPlot(multiome, features="nFeature_RNA", group.by="sample") + geom_hline(yintercept = 2300, color = "black", size=1)

#Check number of cells from each cluster to be removed before filtering 
table(Idents(multiome), multiome$nCount_RNA > 5000 & multiome$nFeature_RNA > 2300) 

#Filter 
multiome_f <- subset(
 x = multiome,
 subset = nCount_RNA > 5000 &
 nFeature_RNA > 2300
   )

DimPlot(multiome_f, reduction = "umap", label =TRUE)
DimPlot(multiome, reduction = "umap", label =TRUE)

multiome <- multiome_f
rm(multiome_f)
```

#Re-run SCT transform with filtered object 
```{r}
multiome[["RNA"]] <- CreateAssayObject(counts = multiome[["SCT"]]$counts)
DefaultAssay(multiome) <- "RNA"
multiome <- SCTransform(multiome, vars.to.regress = c("percent.mt", vars.to.regress = c("percent.mt", "S.Score", "G2M.Score")))
multiome <- RunPCA(multiome)

```

```{r}
#Visualize the significance of each dimension (PC)
DimHeatmap(multiome, dims = 10:15, cells = 500, balanced = TRUE)
ElbowPlot(multiome)


multiome <- FindNeighbors(multiome, dims = 1:13)
multiome <- FindClusters(multiome, resolution = 0.37)
multiome <- RunUMAP(multiome, dims = 1:17, reduction = "pca", seed.use=46)
DimPlot(multiome, reduction = "umap_rna", label =TRUE)



```

#Check previous findings still visible in the reduced dataset
```{r fig.width=10, fig.height=4}
FeaturePlot(multiome, features = c("Lef1", "Tcf7l1"), blend = TRUE, cols = c("lightgrey", "#ff0000", "#00cc00"), order=TRUE,    max.cutoff = "q95")
```



#ATAC-clustering
```{r}
DefaultAssay(multiome) <- "peaks"
multiome <- FindTopFeatures(multiome, min.cutoff = 5)
multiome <- RunTFIDF(multiome)
multiome <- RunSVD(multiome)
```

```{r}
DepthCor(multiome)
```
```{r}
DefaultAssay(multiome) <- "peaks"
multiome <- FindNeighbors(object = multiome, reduction = 'lsi', dims = 2:15)
multiome <- FindClusters(object = multiome, verbose = FALSE, algorithm = 3, resolution= 0.35)
multiome <- RunUMAP(object = multiome, reduction = 'lsi', dims = 2:15)

DimPlot(multiome, reduction = "umap", label =TRUE)

multiome[["umap_atac"]] <- multiome[["umap"]]


DimPlot(multiome, reduction = "umap_atac", label =TRUE)
DimPlot(multiome, reduction = "umap_rna", label =TRUE)
FeaturePlot(multiome, reduction = "umap_atac", features = c("nCount_ATAC", "nFeature_ATAC", "TSS.enrichment"))


```
##Multimodal UMAP
```{r}

multiome <- FindMultiModalNeighbors(
  multiome, 
  reduction.list = list("pca", "lsi"), # Combine RNA (pca) and ATAC (lsi)
  dims.list = list(1:13,
                   2:14)         # Dimensions to use for each
)


multiome <- RunUMAP(
  multiome, 
  nn.name = "weighted.nn",   # Use the combined graph
  reduction.name = "wnn.umap", 
  reduction.key = "wnnUMAP_",
  use.seed=35
)

# Plot the result
DimPlot(multiome, reduction = "wnn.umap", group.by = "sample") + 
  ggtitle("Joint RNA + ATAC UMAP")

multiome[["joint_umap"]] <- multiome[["umap"]]

```

```{r}
setwd("E:/PhD/phd/thesis_figures/Single_Cell_Chapter/multiome")

saveRDS(multiome, file = "multiome_regressed_vars_23012026.rds")
```

